---
title: "DATA 621 HW 1"
author: "Oluwakemi Omotunde"
date: "February 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

DATA 621 - Homework #1
Linear Models with R, 1.1, 1.3, 1.4, 1.5


1. The dataset teengamb concerns a study of teenage gambling in Britain. Make a numerical and graphical summary of the data,commenting on any features that you find intresing. Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data. 

```{r load data}
library(faraway)
data("teengamb")
head(teengamb)
tail(teengamb)
summary(teengamb)
```

Looking at the data, nothing odd jumps out at the immediately. I did notice that one gender is 0, while the other is 1. I'm going to go ahead and sort each of the catagories to see if I can find any interesting things. 

```{r sort}
sort(teengamb$status)
sort(teengamb$income)
sort(teengamb$verbal)
sort(teengamb$gamble)
```
The sorted data did not bring much to light in terms of missing data or outliers. I would like to see how many women and men we have. 

```{r levels}
teengamb$sex <- factor(teengamb$sex)
summary(teengamb$sex)

levels(teengamb$sex) <- c("male", "female") 
summary(teengamb)
```

We will now complete a few different plots to see if we notice anything cool. A few things I'd like to plot are verbal (which I believe is an indication of education) vs. gamble to see if there's a correlation between education and gambling. I would also like to see income vs. gamble and lastly I'd want to see socioeconomic status vs gamble. I am working under the assumption that our outcome is "gamble", while our predictors are "sex, status, income, and verbal". 

```{r plot}
hist(teengamb$gamble, xlab = "GAMBLING EXPENDITURE", main = "") #the look of this histogram is a bit worrisome to me, as it is a bit skewed and I wonder is a linear regression is the right move.
plot(density(teengamb$gamble, na.rm = TRUE), main = "")
plot(sort(teengamb$gamble), ylab = "Sorted Expenditure")
```

When we look at the density plot, we notice that there's a break in our data, indicating that our response is not continuous I'll go back and edit the data. It looks like it may be due to the max being 156. I will remove this datapoint and re-evaluate.


```{r plot gamble}
teengamb$gamble[teengamb$gamble > 90] <- NA
plot(gamble ~ sex, teengamb)
hist(teengamb$gamble, xlab = "GAMBLING EXPENDITURE", main = "")
plot(density(teengamb$gamble, na.rm = TRUE), main = "") #note that the gap is not longer in our density plot
plot(sort(teengamb$gamble), ylab = "Sorted Expenditure")
```

Now let's go back to plot the previously stated and see if we can get some information from this. 

```{r plot gamble2}
plot(teengamb$gamble ~ teengamb$verbal)
plot(teengamb$gamble ~ teengamb$status)
plot(teengamb$gamble ~ teengamb$income)

```

Now that we have looked at the different plots, we will look at our data numerically. 

```{r coefficent}
plot(gamble ~ verbal, teengamb)
abline(lm(gamble ~ verbal, teengamb))
lmod <- lm(gamble ~ verbal, teengamb)
coef(lmod)

plot(gamble ~ status, teengamb)
abline(lm(gamble ~ status, teengamb))
lmod2 <- lm(gamble ~ status, teengamb)
coef(lmod2)

plot(gamble ~ income, teengamb)
abline(lm(gamble ~ income, teengamb))
lmod3 <- lm(gamble ~ income, teengamb)
coef(lmod3)
```
Below are our values from above.
+ Verbal: intercept = 26.13, verbal = -1.46
+ Status: intercept = 11.94, status = .096
+ Income: intercept = -3.84, income = 4.46

From our calculations above, we can see that most of the points do not fall on the best fit line. This leads me to believe that my intial thoughts about this dataset were correct. Due to the skewness of the inital histogram, it looks like a linear regression was not the way to go. 

3. The dataset prostate is from a study on 97 men with prostate cancer who were due to recieve a radical prostatectomy. Make a numerical and graphical summary of the data as in the first question. 

```{r data2}
data(prostate)
head(prostate)
tail(prostate)
summary(prostate)
```

For this question, I'm going to go ahead and skip straight to plotting and explain what I see with the plots. 

```{r factor}
prostate$gleason <- factor(prostate$gleason)
prostate$svi <- factor(prostate$svi)
summary(prostate)
#this dataset did not let me know what each of the categories are.
```

I would like to take a look at the histogram to see if this set is more viable that the one from number 1


```{r hist}
hist(prostate$lcavol, xlab = "Cancer Volume", main = "")
plot(density(prostate$lcavol, na.rm = TRUE), main = "")
plot(sort(prostate$lcavol), ylab = "Sorted Cancer Volume")
```

Our histogram is, again a little funny looking and our density plot is also skew, although not quite as skewwed as the one from problem #1. 
```{r plot cancer}
plot(lcavol ~ lweight, prostate)
abline(lm(lcavol ~ lweight, prostate))
lmod4 <- lm(lcavol ~ lweight, prostate)
coef(lmod4)

plot(lcavol ~ age, prostate)
abline(lm(lcavol ~ age, prostate))
lmod5 <- lm(lcavol ~ age, prostate)
coef(lmod5)

plot(lcavol ~ lcp, prostate)
abline(lm(lcavol ~ lcp, prostate))
lmod6 <- lm(lcavol ~ lcp, prostate)
coef(lmod6)

plot(lcavol ~ lbph, prostate)
abline(lm(lcavol ~ lbph, prostate))
lmod7 <- lm(lcavol ~ lbph, prostate)
coef(lmod7)

plot(lcavol ~ lpsa, prostate)
abline(lm(lcavol ~ lpsa, prostate))
lmod8 <- lm(lcavol ~ lpsa, prostate)
coef(lmod8)
```

From the plots we can see that lpsa and lcp have the most data points that are along our best fit line.

4. The dataset sat comes from a study entitled "Getting What You Pay For: The Debate Over Equity in Public School Expenditures." Make a numerical and graphical summary of the data as in the first question.


```{r data3}
data(sat)
head(sat)
tail(sat)
summary(sat)
```


```{r hist sat}
hist(sat$expend, xlab = "Expenditure", main = "")
plot(density(sat$expend, na.rm = TRUE), main = "")
plot(sort(sat$expend), ylab = "Sorted Expenditure")
```


```{r plot sat}

plot(expend ~ ratio, sat)
abline(lm(expend ~ ratio, sat))
lmod9 <- lm(expend ~ ratio, sat)
coef(lmod9)

plot(expend ~ salary, sat)
abline(lm(expend ~ salary, sat))
lmod10 <- lm(expend ~ salary, sat)
coef(lmod10)

plot(expend ~ takers, sat)
abline(lm(expend ~ takers, sat))
lmod11 <- lm(expend ~ takers, sat)
coef(lmod11)

plot(expend ~ verbal, sat)
abline(lm(expend ~ verbal, sat))
lmod12 <- lm(expend ~ verbal, sat)
coef(lmod12)

plot(expend ~ math, sat)
abline(lm(expend ~ math, sat))
lmod13 <- lm(expend ~ math, sat)
coef(lmod13)

plot(expend ~ total, sat)
abline(lm(expend ~ total, sat))
lmod14 <- lm(expend ~ total, sat)
coef(lmod14)
```

5. The dataset divusa contains data on divorces in the United States from 1920 to 1996. Make a numerical and graphical summary of the data as in the first question.


```{r data4}
data(divusa)
head(divusa)
tail(divusa)
summary(divusa)
```


```{r hist divorce}
hist(divusa$divorce, xlab = "DIVORCE", main = "")
plot(density(divusa$divorce, na.rm = TRUE), main = "")
plot(sort(divusa$divorce), ylab = "Sorted Divorce")
```
Just looking at the plots above, I would not perform a linear regression on this dataset. 

```{r plot divorce}

plot(divorce ~ unemployed, divusa)
abline(lm(divorce ~ unemployed, divusa))
lmod9 <- lm(divorce ~ unemployed, divusa)
coef(lmod9)

plot(divorce ~ marriage, divusa)
abline(lm(divorce ~ marriage, divusa))
lmod10 <- lm(divorce ~ marriage, divusa)
coef(lmod10)

plot(divorce ~ femlab, divusa)
abline(lm(divorce ~ femlab, divusa))
lmod11 <- lm(divorce ~ femlab, divusa)
coef(lmod11)

plot(divorce ~ birth, divusa)
abline(lm(divorce ~ birth, divusa))
lmod12 <- lm(divorce ~ birth, divusa)
coef(lmod12)

plot(divorce ~ military, divusa)
abline(lm(divorce ~ military, divusa))
lmod13 <- lm(divorce ~ military, divusa)
coef(lmod13)
```